---
title: "DDS Assignment 5 - Data: Munging-Merging"
author: "Bruce Granger"
date: "September 28, 2017"
output: html_document
URL: https://github.com/G-Bruce/SMU_DoingDataScience.git
---


### ASSIGNMENT 5 @ GitHub  [CLICK HERE](https://github.com/G-Bruce/SMU_DoingDataScience.git)



```{r Reference, echo=FALSE}
# https://cran.r-project.org/web/packages/stringr/vignettes/regular-expressions.html
```

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
#library(dplyr) is needed for the use of select(), summarise()

library(magrittr)
# library(magrittr) is needed for the use with %>%
```

##### Package::Function used in this exercise:
  
  1) utils
    a) read.table()
    b) write.table()
    c) head()
    d) tail()
    
  2) base
    a) setwd()
    b) summary()
    c) grep()
    d) nrow()
    e) prettyNum()
    f) merge()
    g) names()
  
  3)  dplyr
    a) select()
    b) group_by()
    c) summarise()
    
  2) magrittr
    a) %>% (pipe-operator)

##### Backstory: Your client is expecting a baby soon.  However, he is not sure what to name the child.  Being out of the loop, he hires you to help him figure out recently popular names.  He provides for you raw data in order to help you make a decision

### Data Munging

##### Utilize yob2016.txt for this question. This file is a series of popular children's names born in the year 2016 in the United States.  It consists of three columns with a first name, a gender, and the number of children given that name.  However, the data is raw and will need cleaning to make it tidy and usable. 

  1a) First, import the .txt file into R so you can process it.  Keep in mind this is not a CSV file.  You might have to open the file to see what you're dealing with.  Assign the resulting data frame to an object, df, that consists of three columns with human readable column names for each.

```{r Q_1a, echo=TRUE}
# STORE THE PATH TO FILE VARIABLE INTO THE PATH OBJECT
path<-"/Users/bgranger/Documents/SMU/DDS/SMU_DoingDataScience/Assignment_5"
# MAKE THE PATH THE WORKING (ACTIVE) DIRECTORY
setwd(path)
# READ YOB2016 DATA INTO THE OBJECT DF
df<-read.table("yob2016.txt", sep = ";", header = FALSE, col.names = c("Name","Gender", "Freq"), stringsAsFactors = FALSE)
```

  1b) Display the summary and structure of df.
  
```{r Q_1b, echo=TRUE}
# PRODUCE A SUMMARY OF THE DF OBJECT
summary(df)
# PRODUCE A STRUCTURE (STR) OF THE DF OBJECT
str(df)
# DISPLAY THE NUMBER OF ROWS IN THE DF OBJECT
pre_record_count<-nrow(df)
```

  1c) Your client tells you that there is a problem with the raw file.  One name was entered twice and misspelled.  The client cannot remember which name it is; there are thousands he saw! But he did mention he accidentally put three y's at the end of the name.  Write an R command to figure out which name it is and display it. 
  
```{r Q_1c, echo=TRUE}
# ISOLATE WHERE THE CONDITION ("yyy" OCCURS AT THE END OF A STRING) OCCURS IN THE DF OBJECT AND ASSIGN       TO AN OBJECT CALLED X
x<-df[grep("yyy$",df$Name),]
x
```

  1d) Upon finding the misspelled name, please remove this particular observation, as the client says it's redundant.  Save the remaining dataset as an object: y2016 
  
```{r Q_1d, echo=TRUE}
# FREQUENCY OF INSTANCES THAT END IN "YYY" AND ASSIGN IT TO THE OBJECT yyy_freq
yyy_freq<- df[grep("yyy$",df$Name),]

# SUBSET THE INSTANSTANCE WHERE THE NAME ENDS IN YYY OUT OF THE DATASET INTO AN OBJECT CALLED y2016
y2016<-df[-grep("yyy$",df$Name),]

# NUMBER OF ROWS IN THE y2016 OBJECT
post_record_count<-nrow(y2016)

#RECONCILIATION STEP - PRINT ROW COUTNS
sprintf("The total frequecy represented by the wildcard removal of names ending with 'yyy' was %s.", prettyNum(yyy_freq$Freq, big.mark=","))

sprintf("The total number of rows before the wildcard removal of names ending with 'yyy' was %s, after the removal it is %s.", prettyNum(pre_record_count, big.mark=","), prettyNum(post_record_count, big.mark=","))
```

### Data Merging

##### Utilize The yob2015.txt is similar to yob2016, but contains names, gender, and total children given that name for the year 2015.

  2a) Like 1a, please import the .txt file into R.  Look at the file before you do.  You might have to change some options to import it properly.  Again, please give the dataframe human-readable column names.  Assign the dataframe to y2015.   
  
```{r Q_2a, echo=TRUE}
# READ YOB2015 DATA INTO THE OBJECT yob2015, DEFINE THE SEPERATOR AND ASSIGN COLUMN NAMES
yob2015<-read.table("yob2015.txt", sep = ",", header = FALSE, col.names = c("Name","Gender", "Freq"), stringsAsFactors = FALSE)
```
  
  2b) Display the last ten rows in the dataframe:
  
```{r Q_2b, echo=TRUE}
# DISPAY THE LAST TEN ROWS OF THE yob2015 OBJECT
tail(yob2015, n=10)
```

  2b con't) Describe something you find interesting about these 10 rows.
  
    * Nothing of immediate interest.
    
  2 NOT REQUIRED) - INVESTIGATING THE TOTAL NUMBER OF NAMES BY GENDER
 
```{r Q_2_NOTREQUIRED, echo=TRUE}
# REQUIRES DPLYR PACKAGE
# NOT REQUIRED BY THE ASSIGNMENT

# DISPLAY - YOB2015 DATA - TOTAL COUNT BY GENDER - SUM THE FREQ VALUES
yob2015 %>% select(Gender, Freq) %>% group_by(Gender) %>% summarise(Count=sum(Freq))

# DISPLAY - y2016 DATA - TOTAL COUNT BY GENDER - SUM THE FREQ VALUES
y2016 %>% select(Gender, Freq) %>% group_by(Gender) %>% summarise(Count=sum(Freq))
```
  
  2c) Merge y2016 and y2015 by your Name column; assign it to final.  
  
    - The client only cares about names that have data for both 2016 and 2015; 
    - there should be no NA values in either of your number of children rows after merging.
    
```{r Q_2c, echo=TRUE}
# MERGE THE TWO DATASETS INTO ONE, 'JOIN' WILL BE BASED UPON NAME AND GENDER FIELDS.
# INCLUDING THE GENDER COLUMN IN THE MERGE WILL ENSURE THE MERGE WILL CONSIDER GENDER FOR NAMES SPELLED
#   THE SAME BETWEEN MALE AND FEMALE

final_all<- merge(yob2015, y2016, by=c("Name","Gender"), all = TRUE)
prettyNum(nrow(final_all), big.mark=",")

final_na <- final_all[!complete.cases(final_all),]
prettyNum(nrow(final_na), big.mark=",")

final <- final_all[complete.cases(final_all),]
prettyNum(nrow(final), big.mark=",")
```
  
### Data Summary

##### Utilize your data frame object final for this part. 

  3a) Create a new column called "Total" in final that adds the number of children in 2015 and 2016 together.  In those two years combined, how many people were given popular names?
  
```{r Q_3a, echo=TRUE}
# CREATE A NEW COLUMN CALLED TOTAL, THE CONTENTS OF THE CELL IS THE ADDITION OF THE FREQ.X AND THE FREQ.Y VALUES
final["Total"]<-final$Freq.x + final$Freq.y
```

  3b) Sort the data by Total.  What are the top 10 most popular names? 
  
```{r Q_3b, echo=TRUE}
# STORT THE FINAL DATAFRAME OBJECT BASED UPON THE VALUES IN THE TOTAL COLUMN, DESENDING ORDER
final<- final[order(-final$Total),]
# DISPAY THE FIRST TEN ROWS OF THE FINAL OBJECT
head(final, n=10)
```

  3c) The client is expecting a girl!  Omit boys and give the top 10 most popular girl's names. Assign this to object girl.
  
```{r Q_3c, echo=TRUE}
# SUBSET THE DATAFRAME OBJECT FINAL WHERE GENDER == F (FEMALE) AND ASSIGN IT TO THE OBJECT FINAL
girl <- final[ which(final$Gender =='F'), ]
# STORT THE GIRL DATAFRAME OBJECT BASED UPON THE VALUES IN THE TOTAL COLUMN, DESENDING ORDER
girl<- girl[order(-girl$Total),]
# SLICE THE GIRL DATAFRAME OBJECT TAKING ROWS 1 THROUGH 10 AND ALL COLUMNS
girl<- girl[1:10,]
# DISPLAY THE DATAFRAME OBJECT GIRL
girl
```

  3d) Write these top 10 girl Names and their Totals to a CSV file called itsagirl.csv.  Do not include row labels.  Leave out columns other than Name and Total.

```{r Q_3d, echo=TRUE}
# ASSIGN CHARACTER VECTOR TO myvars, WHICH REPRESENT THE COLUMN NAMES 
myvars <- c("Name","Total")
# EXTRACT/SUBSET COLUMNS INTO THE GIRL DATAFRAME OBJECT  
girl <- girl[myvars]
# DISPLAY THE DATAFRAME OBJECT GIRL
girl
# WRITE THE GIRL OBJECT TO A FILE, OMITTING THE ROW AND COLUMN NAMES
write.table(girl, "itsagirl.csv", sep=",", row.names=F, col.names=F)
```

### Upload to GitHub

##### Push at minimum your RMarkdown for this homework assignment and a Readme file with Codebook to one of your GitHub repositories (you might place this in a Homework repo like last week).  It does not have to be too detailed.  The Codebook should contain a short definition of each object you create.  You are welcome and encouraged to add other files-just make sure you have a description and directions that are helpful for the grader. 

    